Statement of Intent

### 2010-11-16 applications/nsf_fellowship/personal_statement.tex
Finished a draft of my personal statement

\begin{center}
  \textbf{Personal Statement}\\
  Matthew Gardner
\end{center}
\vspace{-4mm}

@@

My time working in the Applied Machine Learning lab has been the defining
experience of my undergraduate education.  I had very close interaction with
several professors and many graduate students who taught me what research was
like.  I also learned a great deal more than was taught in my classes in order
to understand the research that was going on in the lab.  Several times in my
classes the professor had me give part of a lecture that dealt with things I
was actively researching.

While I have tutored other students, mostly informally, throughout all of my
college career, my first formal teaching experience came my first semester as a
masters student.  I was a teaching assistant for two classes that Dr. Seppi
taught: Introduction to Articial Intelligence, and Bayesian Inference.  I had
always enjoyed helping people understand concepts, but I didn't think I would
enjoy the teaching aspect of being a professor until that experience.  I found
that figuring out how to explain complicated concepts simply and seeing the
light of understanding in the students' faces gave me great satisfaction.

I envision my future career as one of research, likely in the domain of natural
language processing, though possibly also in computer vision or other areas of
machine learning.  I have not yet decided between research in industry or
academia; perhaps my career will take me first to industry, ending up as a
professor later in life.

I have spent some time as a teaching assistant, and I
feel that I would enjoy teaching students of all the interesting problems there
are in computer science, getting them excited to do research in the field.  But
what I most want from my career is to find new problems to solve that will
better society, then solve them.  Things like machine translation and
interpretation, or extracting knowledge from large collections of text and
making that knowledge accessible to people, are intriguing problems that I
would love to spend my life's work on.

>>>>

\begin{center}
  \textbf{Personal Statement}\\
  Matthew Gardner
\end{center}

@@

My time working in the Applied Machine Learning lab has been the defining
experience of my undergraduate education.  I had very close interaction with
several professors and many graduate students who taught me what research was
like.  From those that I interacted with, I gained a love of learning new
things and exploring new ideas.  I also learned a great deal more than was
taught in my classes in order to understand the research that was going on in
the lab.  Several times in my classes the professor had me give part of a
lecture that dealt with things I was actively researching.

My undergraduate education would not have been nearly so fulfilling and
enjoyable without my experience working in a research lab.  My understanding of
the ideal college experience has been fundamentally changed through my
research.  If I ever end up as a professor, which I hope to be some day, one of
my goals will be to help all of the undergraduates in my program have the same
kind of experience that I had.  Working closely with professors and graduate
students and conducting independent research were wonderful opportunities that
every undergraduate student should be able to enjoy.

I also had a great internship experience as an undergraduate that helped to
shape my vision of what I wanted from my career.  I worked as an intern on the
TV Ads team at Google during the summer of 2009.  My intern host presented a
project to me that I didn't think was very interesting or very important, so I
proposed a more meaningful, and more difficult, project.  My host supported me
and gave me the help that I needed, and the project turned out to be very
successful.  What I learned from the experience was that while I enjoy writing
software, I find much more satisfication in solving interesting problems that
serve some useful purpose.

As I was nearing the end of my undergraduate career, I found that I had several
research projects that needed just a little more time to be finished.  I wanted
to get into one of the nation's top graduate schools for a PhD, and I thought
that having several publications, instead of hopes of publishing, would give me
a much stronger application.  I also had taken several graduate-level classes
as an undergraduate that wouldn't transfer well to another university, and I
was really enjoying working with Dr. Seppi.  All of these pointed me towards a
masters degree at BYU, so I decided to stay, hoping to finish within a year.

My first semester as a masters student during the summer of 2010 gave me my
first real teaching opportunity.  While I have tutored other students, mostly
informally, throughout all of my college career, I had never done so in a
formal setting.  That semester, however, I was a teaching assistant for two
classes that Dr. Seppi taught: Introduction to Articial Intelligence, and
Bayesian Inference.  I had always enjoyed helping people understand concepts,
but I didn't think I would enjoy the teaching aspect of being a professor until
that experience.  I found that figuring out how to explain complicated concepts
simply and seeing the light of understanding in the students' faces gave me
great satisfaction.

My experience so far as a masters student has served to strengthen my desire to
get a PhD.  I love learning and solving new problems.  I envision my future
career as one of research, likely in the domain of natural language processing,
though possibly also in computer vision or other areas of machine learning.  I
have not yet decided between research in industry or academia.  I enjoy
teaching, but often the resources available in industry allow for more cutting
edge research; the massive amounts of data and extensive computing resources at
Google allow them to tackle problems that are impossible on smaller scales.
Perhaps my career will take me first to industry, ending up as a professor
later in life.  But what I most want from my career is to find new problems to
solve that will better society, then solve them.  Things like machine
translation and interpretation, or extracting knowledge from large collections
of text and making that knowledge accessible to people, are intriguing problems
that I would love to spend my life's work on.
